Trends from 30-Year Observations of Downward Solar Irradiance in Thessaloniki, Greece

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 22:03

<!-- \begin{figure}[h!] -->
<!--     \begin{adjustwidth}{-\extralength}{0cm} -->
<!--         {\centering  -->
<!--             \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly -->
<!--         } -->
<!--     \end{adjustwidth} -->
<!--     \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL} -->
<!-- \end{figure} -->

\FloatBarrier

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

\ \

@@

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki, Greece 
(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`), aiming to identify the long-term 
variability of solar irradiance under different sky conditions.
Under all-sky conditions, there is a positive trend in SDR of 
$`r alltrend`\,\%/y$ (brightening). 
A previous study [@Bais2013] for the period 1993--2011 also reported a
positive trend of $0.33\,\%/y$.
The slight increase in this trend indicates that the brightening of SDR
continues and is likely caused by continuing decreases in aerosol optical
depth and the optical thickness of clouds over the area.
A smaller trend has been found under clear-sky conditions
($`r cleartrend`\,\%/y$), which supports the notion that part of the
brightening is caused by decreasing aerosols.
@Siomos2020 showed that aerosol optical depth over Thessaloniki
was decreasing constantly, at least up until 2018.
The attenuation of SDR by aerosols over Europe has been proposed as
major factor by @Wild2021.
Unfortunately, for this study, aerosol data for the entire period were
not available in order to quantify their effect on SDR. 
The brightening effect on SDR under cloudy-sky conditions
($`r cloudtrend`\,\%/y$) suggests that cloud optical thickness 
is also decreasing during this period.
As long-term data of cloud optical thickness are
also not available for the region, we cannot draw quantitative
conclusions.
<!-- @Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.).  -->
 <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
<!-- We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$. -->

The observed brightening on SDR over Thessaloniki is dependent on
SZA (larger SZAs lead to stronger brightening).
The trend is also dependent on season, with winter showing the 
strongest statistically significant trend of
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
for all- and cloudy-skies, respectively, in contrast to spring and summer.
The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
for all- and cloudy-skies, respectively).
<!-- Our findings are in agreement with other studies for the region. -->
The trend for clear skies is largest in winter
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$) 
and negligible in spring, summer and autumn.

>>>>

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
        }
    \end{adjustwidth}
    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
\end{figure}

@@

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki, Greece 
(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`), aiming to identify the long-term 
variability of solar irradiance under different sky conditions.
Under all-sky conditions, there is a positive trend in SDR of 
$`r alltrend`\,\%/y$ (brightening). 
A previous study [@Bais2013] for the period 1993--2011 also reported a
positive trend of $0.33\,\%/y$.
The slight increase in this trend indicates that the brightening of SDR
continues and is likely caused by continuing decreases in aerosol optical
depth and the optical thickness of clouds over the area.
A smaller trend has been found under clear-sky conditions
($`r cleartrend`\,\%/y$), which supports the notion that part of the
brightening is caused by decreasing aerosols.
@Siomos2020 showed that aerosol optical depth over Thessaloniki
was decreasing constantly, at least up until 2018.
The attenuation of SDR by aerosols over Europe has been proposed as
major factor by @Wild2021.
Unfortunately, for this study, aerosol data for the entire period were
not available in order to quantify their effect on SDR. 
The brightening effect on SDR under cloudy-sky conditions
($`r cloudtrend`\,\%/y$) suggests that cloud optical thickness 
is also decreasing during this period.
As long-term data of cloud optical thickness are
also not available for the region, we cannot draw quantitative
conclusions.

The observed brightening on SDR over Thessaloniki is dependent on
SZA (larger SZAs lead to stronger brightening).
The trend is also dependent on season, with winter showing the 
strongest statistically significant trend of
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
for all- and cloudy-skies, respectively, in contrast to spring and summer.
The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
for all- and cloudy-skies, respectively).
<!-- Our findings are in agreement with other studies for the region. -->
The trend for clear skies is largest in winter
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$) 
and negligible in spring, summer, and autumn.

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 22:03

\FloatBarrier

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

~~

@@

The observed brightening on SDR over Thessaloniki is dependent on SZA
(larger SZAs lead to stronger brightening). The trend is also dependent
on season, with winter showing the strongest statistically significant
trend of \(0.7\) and \(0.76\,\%/y\) for all- and cloudy-skies,
respectively, in contrast to spring and summer. The trends for autumn
are also significant but smaller (\(0.42\) and \(0.19\,\%/y\) for all-
and cloudy-skies, respectively). The trend for clear skies is largest in
winter (\(0.4\,\%/y\)) and negligible in spring, summer and autumn.

>>>>

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
        }
    \end{adjustwidth}
    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
\end{figure}

@@

The observed brightening on SDR over Thessaloniki is dependent on SZA
(larger SZAs lead to stronger brightening). The trend is also dependent
on season, with winter showing the strongest statistically significant
trend of \(0.7\) and \(0.76\,\%/y\) for all- and cloudy-skies,
respectively, in contrast to spring and summer. The trends for autumn
are also significant but smaller (\(0.42\) and \(0.19\,\%/y\) for all-
and cloudy-skies, respectively). The trend for clear skies is largest in
winter (\(0.4\,\%/y\)) and negligible in spring, summer, and autumn.

### 2024-01-02 appendix.tex
Commit 2024-01-01 18:12
Commit 2024-01-02 22:03

%% From raw MDPI template
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}

@@

    }
    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for 1993 -- 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
\end{figure}

@@

    }
    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 -- 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
\end{figure}

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
        \caption{Long term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$).
               Cases where $p<0.005$ or with less than $85$ observations may be
               missing from view.}\label{fig:SZAtrendSeason}
    \end{adjustwidth}
\end{figure}

>>>>

%% From raw MDPI template
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}
\label{appendix}

@@

    }
    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for 1993--2023. The black line shows the long-term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
\end{figure}

@@

    }
    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993--2023. The black line shows the long-term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
\end{figure}

\begin{figure}[h!]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
    \end{adjustwidth}
    \caption{Long-term trends of SDR as a function of SZA separately form morning
           and afternoon periods, by season (rows of plots) for the three sky
           conditions (columns of plots).
           Solid shapes represent statistical significant trends ($p<0.005$), (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.
           Cases where $p<0.005$ or with less than $85$ observations may be
           missing from view.}\label{fig:SZAtrendSeason}
\end{figure}
